---
output: github_document
---

```{r setup, include = F}
# devtools::install_github("Hemken/Statamarkdown")
library(Statamarkdown)
```
# Hansen DWI Replication

**Directions:** Download `hansen_dwi.dta` from GitHub at the following address. Note these data are not exactly the same as his because of confidentiality issues (so he couldn’t share all of it).

https://github.com/scunning1975/causal-inference-class/raw/master/hansen_dwi

The outcome variable is `recidivism` which is measuring whether the person showed back up in the data within 4 months. Use this data to answer the following questions.

```{r load-data}
## hansen.R --------------------------------------------------------------------
## Kyle Butts, CU Boulder Economics
## 
## replicate figures and tables in Hansen 2015 AER

library(fixest)
library(ggplot2)
library(rdrobust)
library(rddensity)
library(binsreg)

# load the data from github
df <- haven::read_dta("https://github.com/scunning1975/causal-inference-class/raw/master/hansen_dwi.dta")
```

1. We will only focus on the 0.08 BAC cutoff; not the 0.15 cutoff. Take the following steps.
  a. Create a treatment variable (`dui`) equaling 1 if `bac1 >= 0.08` and 0 otherwise in your do/R file.

```{r create-dui}
# 1.a. create dui treatment variable for bac1>=0.08
df$dui = (df$bac1 > 0.08)
```

  b. Store `bac1` into `bac1_orig` and then center the `bac1` variable, i.e. subtract $0.08$ so that the cutoff is now zero. 

```{r recenter-bac1}
# 1.b. Re-center our running variable at bac1=0.08
df$bac1_orig = df$bac1
df$bac1 = df$bac1 - 0.08
```

  c. Replicate Hansen’s Figure 1 examining whether there is any evidence for manipulation on the running variable. Produce a raw histogram using `bac1`, then use the density test in Cattaneo, Titunik and Farrell’s `rddensity` package. Can you find any evidence for manipulation? What about heaping?

```{r manipulation-heaping-tests}
# 1.c. Find evidence for manipulation or heaping using histograms
ggplot(df) + 
  geom_histogram(
    aes(x = bac1), binwidth = 0.001,
    alpha = 0.8, color = "steelblue"
  ) + 
  labs(
    x = "Blood Alcohol Content",
    y = "Frequency",
    title = "Replicating Figure 1 of Hansen AER 2015"
  ) + 
  theme_bw()

# Use rddensity from Cattnaeo, Titunik and Farrell papers
rddensity::rddensity(X = df$bac1, c = 0.08) |> summary()
```

2. We are going to test for manipulation around the cutoff (following Table 2 Panel A). Run RD regressions using a local-linear estimator on `white`, `male`, age (`aged`) and accident (`acc`) as dependent variables. Are the covariates balanced at the cutoff? Use data in `bac1_orig` 0.03 to 0.13 (or `bac1` in -0.05 to 0.05). Check if the results are robust to a more narrow bandwidth of `bac1_orig` in 0.055 to 0.105. 

```{r covariate-balance}
# 2. Are the covariates balanced at the cutoff? 
# Use two separate bandwidths (0.03 to 0.13; 0.055 to 0.105)
# yi = Xi′γ + α1 DUIi + α2 BACi + α3 BACi × DUIi + ui
feols(
  c(white, male, acc, aged) ~ dui + bac1 + i(dui, bac1), 
  df[df$bac1_orig >= 0.03 & df$bac1_orig <= 0.13, ], vcov = "hc1"
) |> 
  etable()

feols(
  c(white, male, acc, aged) ~ dui + bac1 + i(dui, bac1), 
  df[df$bac1_orig >= 0.055 & df$bac1_orig <= 0.105, ], vcov = "hc1"
) |> 
  etable()
```

3. Now, we turn our main result, estimating the effect of getting a DUI on recidivism (`recid`).
  a. Run an RD estimate using the `rdrobust` command (from the `rdrobust` package in R)

```{r rdrobust}
# 3. Estimate RD of DUI on Recidivism
rdrobust(
  y = df$recidivism, x = df$bac1, c = 0
)
```

  b. Like all RD applications, you need to include a plot of the underlying data. Plot the RD estimator using the `rdplot` command (from the `rdrobust` package in R).

```{r rdplot}
rdplot(
  y = df$recidivism, x = df$bac1, c = 0
)
```

4. Repeat but drop units in the close vicinity of 0.08 (0.079-0.081) (i.e., the "donut hole" regression). Do the results stay the same?

```{r donut}
df_donut <- df[df$bac1_orig <= 0.79 | df$bac1_orig >= 0.081, ]
rdrobust(
  y = df_donut$recidivism, x = df_donut$bac1, c = 0
)
```

